---
title: "Mini-Project 01 - Assessing the Impact of SFFA on Campus Diversity One-Year Later"
author: "JanielAve"
format: html
editor:
  mode: source
execute:
  echo: false
  message: false
  warning: false
---

```{r setup, include=FALSE}

library(tidyverse)
library(stringr)
library(skimr)
library(scales)

## Introduction

** In this mini-project, I intend to demonstrate
my skills at Data Summarization/Exploratory Analysis/Table Formatting in an analysis of Official Statistics/Demography data.**
  
  ## Data Acquisition and Preparation
  cquire_ipeds_data <- function(start_year=2010, end_year=2024){
    library(tidyverse)
    library(glue)
    
    data_dir <- file.path("data", "mp01")
    
    if(!dir.exists(data_dir)){
      dir.create(data_dir, showWarnings=FALSE, recursive=TRUE)
    }
    
    YEARS <- seq(start_year, end_year)
    
    EFA_ALL <- map(YEARS, function(yy){
      if(yy <= 2022){
        ef_url <- glue("https://nces.ed.gov/ipeds/datacenter/data/EF{yy}A.zip")
        
      } else {
        ef_url <- glue("https://nces.ed.gov/ipeds/data-generator?year={yy}&tableName=EF{yy}A&HasRV=0&type=csv")
      }
      
      ef_file <- file.path(data_dir, glue("ef{yy}a.csv.zip"))
      
      if(!file.exists(ef_file)){
        message(glue("Downloading Enrollment Data for {yy} from {ef_url}"))
        download.file(ef_url, destfile = ef_file, quiet=TRUE, mode="wb")    
      }
      
      read_csv(ef_file, 
               show_col_types=FALSE) |>
        mutate(year = yy, 
               # American Indian or Alaskan Native
               enrollment_m_aian = EFAIANM, 
               enrollment_f_aian = EFAIANW, 
               # Asian
               enrollment_m_asia = EFASIAM, 
               enrollment_f_asia = EFASIAW, 
               # Black or African-American, 
               enrollment_m_bkaa = EFBKAAM, 
               enrollment_f_bkaa = EFBKAAW, 
               # Hispanic 
               enrollment_m_hisp = EFHISPM, 
               enrollment_f_hisp = EFHISPW, 
               # Native Hawaiian or Other Pacific Islander 
               enrollment_m_nhpi = EFNHPIM, 
               enrollment_f_nhpi = EFNHPIW, 
               # White
               enrollment_m_whit = EFWHITM, 
               enrollment_f_whit = EFWHITW, 
               # Two or More Races
               enrollment_m_2mor = EF2MORM, 
               enrollment_f_2mor = EF2MORW, 
               # Unknown / Undisclosed Race
               enrollment_m_unkn = EFUNKNM, 
               enrollment_f_unkn = EFUNKNW, 
               # US Non-Resident
               enrollment_m_nral = EFNRALM, 
               enrollment_f_nral = EFNRALW, 
        ) |> filter(
          (EFALEVEL %in% c(2, 12)) | (LINE %in% c(1, 15))
          # Per 2024 Data Dictionary, 
          # - EFALEVEL 2 = undergrad 
          # - EFALELVE 12 = grad
          # - Line 1 = first year first time full-time undergrad
          # - Line 15 = first year first time part-time undergrad
        ) |> mutate(level = case_when(
          EFALEVEL == 2 ~ "all undergrad", 
          EFALEVEL == 12 ~ "all graduate",
          LINE %in% c(1, 15) ~ "first year undergrad"
        )
        ) |>
        select(institution_id = UNITID, 
               year, 
               level,
               starts_with("enrollment_")) |>
        group_by(institution_id, 
                 year, 
                 level) |>
        summarize(across(starts_with("enrollment_"), sum), 
                  .groups = "drop")
      
    }) |> bind_rows()
    
    DESC_ALL <- map(YEARS, function(yy){
      if(yy <= 2022){
        hd_url <- glue("https://nces.ed.gov/ipeds/datacenter/data/HD{yy}.zip")
        
      } else {
        hd_url <- glue("https://nces.ed.gov/ipeds/data-generator?year={yy}&tableName=HD{yy}&HasRV=0&type=csv")
      }
      
      hd_file <- file.path(data_dir, glue("hd{yy}.csv.zip"))
      
      if(!file.exists(hd_file)){
        message(glue("Downloading Institutional Descriptions for {yy} from {hd_url}"))
        download.file(hd_url, destfile = hd_file, quiet=TRUE, mode="wb")    
      }
      
      suppressWarnings(
        read_csv(hd_file, 
                 show_col_types=FALSE, 
                 locale=locale(encoding=if_else(yy==2024, "utf-8", "windows-1252"))) |>
          mutate(year = yy, 
                 INSTNM) |> 
          select(institution_id = UNITID, 
                 institution_name = INSTNM, 
                 state = STABBR, 
                 year)
      )
      
    }) |> bind_rows()
    
    inner_join(EFA_ALL, 
               DESC_ALL, 
               join_by(institution_id == institution_id, 
                       year == year))
  }

IPEDS <- acquire_ipeds_data()


Code related to Data Acquisition and Preparation goes here.

Remember that code blocks look like this: 
  
  ```{r}

## Exploratory Analysis


library(dplyr)
library(stringr)
IPEDS <- IPEDS |>
  mutate(is_cuny = str_detect(institution_name, "CUNY"),
         is_suny = str_detect(institution_name, "SUNY"),
         is_calpublic = str_detect(institution_name, "University of California") |
           str_detect(institution_name, "California State University"))

cal_public_list <- IPEDS |>
  filter(is_calpublic == TRUE) |>
  select(institution_name)
print(cal_public_list)
is_suny_list <- IPEDS |>
  filter(is_suny == TRUE) |>
  select(institution_name)
print(is_suny_list)
is_cuny_list <- IPEDS |>
  filter(is_cuny == TRUE) |>
  select(institution_name)
print(is_cuny_list)

glimpse(IPEDS)
summary(IPEDS)

install.packages("skimr")
library(skimr)
skim(IPEDS)

**#1. How many distinct institutions appear in this data set?**
  IPEDS |> 
  summarize(distinct_count = n_distinct(institution_name))

**#2 How many graduate students were enrolled at Baruch & Farmingdale in 2024?**
  IPEDS |> 
  filter(year == 2024, 
         str_detect(institution_name, "Baruch"),
         level == "all graduate") |>
  mutate(total_grad = rowSums(across(starts_with("enrollment_")))) |> 
  select(institution_name, total_grad)

IPEDS |> 
  filter(year == 2024, 
         str_detect(institution_name, "Farmingdale"),
         level == "all graduate") |>
  mutate(total_grad = rowSums(across(starts_with("enrollment_")))) |> 
  select(institution_name, total_grad)
**#3 How many total students were enrolled at Baruch and Farmingdale in 2024?**
  IPEDS |> 
  filter(year == 2024, 
         str_detect(institution_name, "Baruch"), 
         level %in% c("all undergrad", "all graduate")) |> 
  mutate(row_total = rowSums(across(starts_with("enrollment_")))) |> 
  summarize(baruch_total_2024 = sum(row_total))
IPEDS |> 
  filter(year == 2024, 
         str_detect(institution_name, "Farmingdale"), 
         level %in% c("all undergrad", "all graduate")) |> 
  mutate(row_total = rowSums(across(starts_with("enrollment_")))) |> 
  summarize(baruch_total_2024 = sum(row_total))
**#4 Which institution had the highest number of enrolled female students in 2019?**
  IPEDS |> 
  filter(year == 2019, 
         level %in% c("all undergrad", "all graduate")) |> 
  mutate(female_in_row = rowSums(across(starts_with("enrollment_f_")))) |> 
  group_by(institution_name) |> 
  summarize(total_female = sum(female_in_row), .groups = "drop") |> 
  slice_max(total_female, n = 1)
**#5 Which institution with over 1000 total students admitted the highest proportion 
  #of Native Hawaiian or Pacific Islanders first year undergraduates in 2024?**
  library(scales)
IPEDS |> 
  mutate(total_students = rowSums(across(starts_with("enrollment_")))) |> 
  filter(year == 2024, total_students > 1000, level == "first year undergrad") |> 
  mutate(
    nhpi_total = enrollment_m_nhpi + enrollment_f_nhpi,
    nhpi_prop = nhpi_total / total_students
  ) |> 
  slice_max(nhpi_prop, n = 1) |> 
  select(institution_name, nhpi_prop)
percent(0.0300, accuracy=0.01)

## Final Insights and Deliverable
#Task #6
**#1 Which 5 states had the highest number of graduate students across all institutions
  #located in that state?**
  install.packages("gt")
library(gt)

IPEDS |>
  filter(level == "all graduate") |>
  mutate(grad_total = rowSums(across(starts_with("enrollment_")))) |>
  group_by(state) |>
  summarize(total_grads = sum(grad_total)) |>
  slice_max(total_grads, n = 5) |>
  gt() |>
  fmt_number(columns = total_grads, decimals = 0) |>
  cols_label(state = "State", total_grads = "Total Graduate Students")

**#2 In 2024, how many first year undergraduates students were enrolled at CUNY/SUNY colleges
  #and which colleges did they attend? Report both absolute enrollment numbers and percent
  #of total first year undergraduates? **
  
  library(dplyr)
library(gt)
library(stringr)
#Filter for 2024 First-Year students at CUNY/SUNY schools

IPEDS <- IPEDS |>
  mutate(
    is_suny = str_detect(institution_name, fixed("SUNY", ignore_case = TRUE)) | 
      str_detect(institution_name, fixed("Farmingdale", ignore_case = TRUE)),
    is_cuny = str_detect(institution_name, fixed("CUNY", ignore_case = TRUE))
  )

# Filter First-Year students
cuny_suny_2024 <- IPEDS |>
  filter(year == 2024, 
         (is_cuny | is_suny), 
         level == "first year undergrad") |>
  mutate(first_year_count = rowSums(across(starts_with("enrollment_"))))

#combined total
total_combined_fy <- sum(cuny_suny_2024$first_year_count, na.rm = TRUE)

#table
cuny_suny_2024 |>
  mutate(
    pct_of_total = first_year_count / total_combined_fy,
    system = if_else(is_cuny, "CUNY", "SUNY")
  ) |>
  select(system, institution_name, first_year_count, pct_of_total) |>
  arrange(system, desc(first_year_count)) |>
  gt() |>
  tab_header(title = "CUNY and SUNY First-Year Enrollment (2024)") |>
  fmt_number(columns = first_year_count, decimals = 0) |>
  fmt_percent(columns = pct_of_total, decimals = 1)

**#3 How has Baruch's/Farmingdale's total undergraduate enrollment changed over the
  #study period? Report both enrollment numbers and percent change year over year.**
  
  library(dplyr)
library(stringr)
library(gt)

IPEDS |>
  filter(str_detect(institution_name, "Baruch") | str_detect(institution_name, "Farmingdale"), 
         level == "all undergrad") |>
  mutate(total_ungr = rowSums(across(starts_with("enrollment_")))) |>
  # Group by institution
  group_by(institution_name) |>
  arrange(year, .by_group = TRUE) |>
  mutate(prev_year_enroll = lag(total_ungr),
         pct_change = (total_ungr - prev_year_enroll) / prev_year_enroll) |>
  ungroup() |>
  # table formatting
  select(institution_name, year, total_ungr, pct_change) |>
  gt() |>
  tab_header(
    title = "Undergraduate Enrollment Trends",
    subtitle = "Baruch College vs. Farmingdale State College"
  ) |>
  fmt_number(columns = total_ungr, decimals = 0) |>
  fmt_percent(columns = pct_change, decimals = 1) |>
  cols_label(
    institution_name = "Institution",
    year = "Year", 
    total_ungr = "Enrollment", 
    pct_change = "YoY Change" )
**#4 At what 5 institutions did the fraction of white students decrease the most
  #over the period from 2010 to 2020?**
  IPEDS |>
  #filtering for the years and levels
  filter(year %in% c(2010, 2020), 
         level %in% c("all undergrad", "all graduate")) |>
  #totals
  mutate(total_in_row = rowSums(across(starts_with("enrollment_"))),
         white_in_row = enrollment_m_whit + enrollment_f_whit) |>
  #Group by
  group_by(institution_name, year) |>
  summarize(total = sum(total_in_row),
            white_total = sum(white_in_row), 
            .groups = "drop") |>
  #calculation
  filter(total > 1000) |>
  mutate(white_frac = white_total / total) |>
  select(institution_name, year, white_frac) |>
  pivot_wider(names_from = year, names_prefix = "yr_", values_from = white_frac) |>
  filter(!is.na(yr_2010) & !is.na(yr_2020)) |>
  mutate(diff = yr_2020 - yr_2010) |>
  slice_min(diff, n = 5) |> 
  gt() |>
  tab_header(title = "Largest Decrease in Fraction of White Students (2010-2020)",
             subtitle = "Combined Undergraduate and Graduate Enrollment") |>
  fmt_percent(columns = c(yr_2010, yr_2020, diff), decimals = 1) |>
  cols_label(institution_name = "Institution", 
             yr_2010 = "2010 %", 
             yr_2020 = "2020 %", 
             diff = "Total Change")
**#5 In which 3 states did the fraction of female undergraduates increase the most
  #over the period from 2010 to 2024?**
  
  IPEDS |>
  filter(year %in% c(2010, 2024), level == "all undergrad") |>
  mutate(total = rowSums(across(starts_with("enrollment_"))),
         female_total = rowSums(across(starts_with("enrollment_f_"))),
         female_frac = female_total / total) |>
  group_by(state, year) |>
  summarize(state_female_frac = mean(female_frac, na.rm = TRUE), .groups = "drop") |>
  pivot_wider(names_from = year, names_prefix = "yr_", values_from = state_female_frac) |>
  mutate(increase = yr_2024 - yr_2010) |>
  slice_max(increase, n = 3) |>
  gt() |>
  fmt_percent(columns = c(yr_2010, yr_2024, increase), decimals = 1) |>
  cols_label(state = "State", yr_2010 = "2010 % Female", yr_2024 = "2024 % Female", increase = "Growth")



Code related to the final deliverable of the assignment goes here. 

#for oped
# 2023 vs 2024 First-Year Comparison

library(dplyr)
library(gt)
fy_comparison_table <- IPEDS |>
  filter(year %in% c(2023, 2024), 
         str_detect(institution_name, "Baruch|Farmingdale"),
         level == "first year undergrad") |>
  mutate(total_fy = rowSums(across(starts_with("enrollment_")))) |>
  select(institution_name, year, total_fy) |>
  pivot_wider(names_from = year, values_from = total_fy, names_prefix = "fy_") |>
  mutate(abs_change = fy_2024 - fy_2023,
         pct_change = abs_change / fy_2023)
fy_comparison_table |>
  gt() |>
  tab_header(
    title = "Incoming First-Year Class Comparison",
    subtitle = "Fall 2023 vs. Fall 2024 Enrollment"
  ) |>
  fmt_number(columns = c(fy_2023, fy_2024, abs_change), decimals = 0) |>
  fmt_percent(columns = pct_change, decimals = 1) |>
  cols_label(
    institution_name = "College",
    fy_2023 = "2023 Class",
    fy_2024 = "2024 Class",
    abs_change = "Difference",
    pct_change = "% Change"
  )

#Demographic Trend( 2023 vs 2024)

library(dplyr)
library(tidyr)
library(stringr)
library(gt)

#Calculation
all_groups_trend <- IPEDS |>
  filter(year %in% c(2023, 2024), 
         str_detect(institution_name, "Baruch|Farmingdale"),
         level %in% c("all undergrad", "all graduate")) |>
  # Sum genders
  mutate(
    total = rowSums(across(starts_with("enrollment_"))),
    aian  = enrollment_m_aian + enrollment_f_aian,
    asia  = enrollment_m_asia + enrollment_f_asia,
    bkaa  = enrollment_m_bkaa + enrollment_f_bkaa,
    hisp  = enrollment_m_hisp + enrollment_f_hisp,
    nhpi  = enrollment_m_nhpi + enrollment_f_nhpi,
    whit  = enrollment_m_whit + enrollment_f_whit,
    twom  = enrollment_m_2mor + enrollment_f_2mor
  ) |>
  group_by(institution_name, year) |>
  summarize(across(c(total, aian, asia, bkaa, hisp, nhpi, whit, twom), sum), .groups = "drop") |>
  # Calculation
  mutate(across(c(aian, asia, bkaa, hisp, nhpi, whit, twom), ~ .x / total, .names = "{.col}_pct")) |>
  select(institution_name, year, ends_with("_pct")) |>
  pivot_wider(names_from = year, values_from = ends_with("_pct"), names_glue = "{.value}_{year}") |>
  # Change Calculation
  mutate(
    hisp_change = hisp_pct_2024 - hisp_pct_2023,
    whit_change = whit_pct_2024 - whit_pct_2023,
    asia_change = asia_pct_2024 - asia_pct_2023,
    bkaa_change = bkaa_pct_2024 - bkaa_pct_2023
  )

#Table
all_groups_trend |>
  gt() |>
  tab_header(
    title = "Post-SFFA Demographic Shift (2023-2024)",
    subtitle = "Year-Over-Year Change in Ethnic Composition"
  ) |>
  # Grouping
  tab_spanner(label = "Hispanic", columns = contains("hisp")) |>
  tab_spanner(label = "White", columns = contains("whit")) |>
  tab_spanner(label = "Asian", columns = contains("asia")) |>
  tab_spanner(label = "Black", columns = contains("bkaa")) |>
  fmt_percent(columns = -institution_name, decimals = 1) |>
  cols_label(
    institution_name = "College",
    hisp_pct_2023 = "'23", hisp_pct_2024 = "'24", hisp_change = "Δ",
    whit_pct_2023 = "'23", whit_pct_2024 = "'24", whit_change = "Δ",
    asia_pct_2023 = "'23", asia_pct_2024 = "'24", asia_change = "Δ",
    bkaa_pct_2023 = "'23", bkaa_pct_2024 = "'24", bkaa_change = "Δ"
  ) |>
  data_color(
    columns = ends_with("_change"),
    fn = scales::col_numeric(palette = c("red", "white", "blue"), domain = c(-0.05, 0.05))
  )

**OP-ED:**
  **Intro:**
  On June 29, 2023, The US Supreme Court changed higher education in the United States of America. The outcome of the case Students for Fair Admissions v. Harvard shifted the way admissions will be performed. The outcome of the case ended an era in which race-conscious admissions practices were executed by higher education organizations. Have our schools changed after no longer having diversity protected at the federal level. We shall analyze the admissions data from 2024 and earlier years to determine any diversity changes at Farmingdale State College and Baruch College, two different environmental settings, but both with the same results. Diversity is still of utmost importance to both our educational organizations.
Our Definition of Diversity: 
  Diversity to us is the ability to provide the opportunity of education to individuals from any background which includes any race, any ethnicity, any gender, and any sexual orientation. Having diversity helps us create a campus and environment which brings diverse viewpoints and ideas in order to grow as an educational organization. Having a diverse environment prepares us all for a world filled with different cultural backgrounds and ideas. 
**Organization Stats:**
  Baruch College:
  CUNY Baruch College is located in the greatest city on Earth, NYC, in 2024 we had a total of 3,585 Graduate students. The total number of students enrolled at Baruch was 20,081. Now for the important number which is the admissions of 2024, our first year undergrad admissions were 2,532 which are welcomed to our community. Our year to year change between 2023-2024 was from 1.2% to 2.5%, therefore we saw growth in our enrollment. Unfortunately, we did see a decrease in our incoming first year students between those years, we saw a drop of -3.1% which is not major, but we must keep monitoring in future years.
Farmingdale State College:
  SUNY Farmingdale State College is located on Long Island in a more suburban area, at Farmingdale we had 32 graduate students, which makes sense because until recently we added graduate programs to our curriculum. The total number of students enrolled at Farmingdale State College was 10,039. The total new undergrad admissions was 1,822. Our year to year change from 2023-2024 was from 2.6% to 5.2% which is an amazing growth amount. Unlike Baruch, at Farmingdale we saw a good amount of growth between the same years when it comes to incoming first year students, the growth was 8.7%. 

**Demographics Changes Between 2023-2024:**
  For Baruch College we only saw a negative change in one of our ethnic groups. That ethnic group is the Asian demographic but it was a -0.2% drop, all other groups stayed the same between both years or increased a bit. For Farmingdale State College we did see a bigger drop in one ethnic group. Unfortunately, we saw a -3.3% shift in the White demographic, but we did see increases in all other demographic ethnic groups. For long term diversity trends in both institutions we do see decreases in the White students fractions, we must understand why this is occurring. For other demographic groups we do see continued increases therefore diversity continues to be part of both our institutions despite the choices of the Supreme Court.

**AI USAGE STATEMENT:**
  Gemini was used to help write the code in this project, but all non-code text was generated without the use of any Generative AI tools. Additionally, Gemini was used to provide additional background information on the topic and to brainstorm ideas for the final open-ended prompt.


------------------------------------------------------------------------
  
  This work ©2025 by JanielAve was initially prepared as a Mini-Project for
STA 9750 at Baruch College. More details about this course can be found at
[the course site](https://michael-weylandt.com/STA9750) and instructions for
this assignment can be found at 
[MP #01](https://michael-weylandt.com/STA9750/mini/mini01.html)
  
  

